---
title: "02_1_prepare_flanders_input"
output: html_document
date: "2026-08-28"
---

02_1_prepare_flanders_inputs.R 

################################################################################

2.1. Prepare Flanders input file

################################################################################

Aim: Prepare the flanders input table and format files for flanders input 

```{r}
library(tidyverse)
library(data.table)
# Set out path 
pops_lustre_path <- "genotyping/analysis/eQTL/"
outpath <- paste0(pops_lustre_path, "input_data/flanders/")
interval_path <- "interval_rna/" 
```

We will fine-map and colocalize the following summary stats
# eQTL
1. POPs2 main effect eQTL 
2. Interval main effect eQTL from full cohort (to determine whether the colocs with pregnancy traits could have been identified in any healthy cohort)
# GWAS
3. GDM GWAS
4. Pre-eclampsia GWAS (FinnGen)
5. Pre-eclampsia GWAS (Meta analysis)
6. sPTB GWAS
7. Gestational duration GWAS

# 0. Make input parameters for flanders
```{r}
# make the input parameters file for Flanders
columns.needed <- c("input", # Path and file name of the GWAS summary statistics file to fine-map.
                    "study_id", # Identifier for the GWAS summary statistics file; used to identify results coming from the same GWAS.       
                    "chr_lab", #  Name of the chromosome column in the GWAS summary statistics.  ## NOTE: X chr must be coded as 23
                    "pos_lab", # Name of the genomic position column in the GWAS summary statistics.     
                    "rsid_lab", # Name of the rsID or other SNP identifier column.      
                    "a1_lab", # Name of the effect allele column.  
                    "a0_lab", # Name of the non-effect allele column.          
                    "freq_lab", # Name of the effect allele frequency column. If not available, set to NA and it will be calculated from the PLINK bfiles.
                    "n_lab", # Name of the sample size column. If not available, set to NA and it will be approximated using MAF and SE.     
                    "effect_lab", # Name of the effect size column (e.g., beta or odds ratio).        
                    "se_lab", # Name of the effect standard error column.                                  
                    "pvalue_lab", #  Name of the p-value column; can be nominal or log10-transformed.     
                    "type", # Trait type: 'quant' for quantitative traits or 'cc' for case–control traits.       
                    "sdY", # For quantitative traits: population standard deviation of the trait (1 if standardised). If NA, will be approximated. NA for binary traits.
                    "s", # For binary traits: proportion of samples that are cases. NA for quantitative traits.   
                    "grch", # Genome build of the GWAS summary statistics (37 or 38).                              
                    "p_thresh1", # Nominal p-value threshold to retain significant association regions (suggested: 5e-08).      
                    "p_thresh2", # Nominal p-value threshold to define borders of association regions (suggested: 1e-05).      
                    "hole", # Minimum base-pair distance to separate independent loci (suggested: 250000).                          
                    "bfile", # Path and file name prefix to PLINK binary files (.bed/.bim/.fam) used for LD computation.     
                    "maf", # Minor allele frequency threshold; variants below this value are excluded.   
                    "is_molQTL", # Boolean flag indicating whether the trait is molecular QTL (e.g., eQTL, pQTL).            
                    "key", # When is_molQTL=TRUE, column identifying the traits (e.g., genes, proteins). Set to NA if FALSE.   
                    "cs_thresh", # Posterior inclusion probability (PIP) threshold for defining credible sets (e.g., 0.99).   
                    "grch_bfile", #  Genome build of LD reference PLINK files (should match grch if no liftover is used).   
                    "process_bfile", # Flag to indicate whether the LD reference PLINK files need to be lifted over and/or processed by the pipeline. Set to FALSE to reuse pre-processed files (optional)
                    "per_gene_p1", # Added by HGI: .tsv or .csv file of per-gene pvalue thresholds (i.e. from eigenMT)
                    "per_gene_p2" # Added by HGI: .tsv or .csv file of per-gene pvalue thresholds (i.e. from eigenMT)
                    )

input.info <- matrix(NA, ncol=28, nrow=7, dimnames = list(seq(1, 7), columns.needed))

```

# 1. POPs2 main effect eQTL 
Make input row
```{r}
# make row for input table 
input.info[1,"input"] <- paste0(outpath, "POPS2_main_effect_input/POPS2_flanders_main_effects.txt")
input.info[1,"sdY"] <- paste0(outpath, "POPS2_main_effect_input/POPS2_gene_sdY.txt")
input.info[1,"per_gene_p1"] <- paste0(outpath, "POPS2_main_effect_input/eqtl_pval_thresh1_per_gene.tsv")
input.info[1,"per_gene_p2"] <- paste0(outpath, "POPS2_main_effect_input/eqtl_pval_thresh2_per_gene.tsv")
input.info[1, "rsid_lab"] <- "snps" # rsID or SNP identifier column name | String | Mandatory
input.info[1, "freq_lab"] <- "MAF" # Effect allele frequency column name | String or NA | Optional
input.info[1, "effect_lab"] <- "beta" # Effect size column (e.g., beta or OR) | String | Mandatory
input.info[1, "se_lab"] <- "se" # Standard error column | String | Mandatory
input.info[1, "pvalue_lab"] <- "pvalue" # P-value column (nominal or log10) | String | Mandatory
input.info[1, "key"] <- 	"gene" # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[1, "type"] <- "quant" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[1, "is_molQTL"] <- 	TRUE
input.info[1, "grch"] <- 38 # Genome build (37 or 38) | Integer | Mandatory
input.info[1, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[1, "maf"] <- 	0.01 # Minor allele frequency threshold | Numeric (0–1) | Mandatory
input.info[1, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[1, "grch_bfile"] <- 38 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[1, "bfile"] <- "genotyping/data/imputed_genotyping_dataset_for_eQTL/POPS2_genotyping_imputed_eQTL_rsid" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory # For Sanger users - this is from POPs2 dir
input.info[1, "chr_lab"] <- "chr" # Chromosome column name in GWAS | String | Mandatory
input.info[1, "pos_lab"] <- "SNPpos" # Genomic position column name in GWAS | String | Mandatory
input.info[1, "a1_lab"] <- "A1" # Effect allele column name | String | Mandatory
input.info[1, "a0_lab"] <- "A2" # Non-effect allele column name | String | Mandatory
input.info[1, "n_lab"] <- "nsamples" # Sample size column name | String or NA | Optional
input.info[1, "study_id"] <- "POPs2_bulk_all" # Identifier for the GWAS study | String | Mandatory
input.info[1,"p_thresh1"] <- 1.00E-05 # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory - I won't be using this anyway 
input.info[1,"p_thresh2"] <- 1.00E-04 # nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory - I won't be using this anyway 
input.info[1, "process_bfile"] <- TRUE
```

Make input files
```{r}
# main effects
main_effects_pops <- readRDS(paste0(pops_lustre_path, "output_data/main_effects/ciseqtl_all.rds"))
# read in p-value thresholds per gene 
pval_thresholds_pops <- read.delim(paste0(pops_lustre_path, "output_data/eigenMT/nominal_pval_thresholds.txt"))
# maf df
maf_df_pops <- read.table(paste0(pops_lustre_path, "input_data/eQTL_maf.frq"), header = TRUE)
# eQTL count matrix 
gene_exp_pops <- readRDS(paste0(pops_lustre_path, "input_data/eQTL_counts.rds"))

# make input files for flanders 
##### Results ##### 
# update the results df to contain the columns needed for flanders
flanders_input <- main_effects_pops %>% left_join(maf_df_pops %>% 
                                                      # update the maf_df to have the correct format for the non-rsid snps
                                               mutate(SNP = ifelse(!grepl("rs", SNP), gsub(":", ".", SNP), SNP),
                                                      # calculate n samples 
                                                      nsamples = NCHROBS/2) %>%
                                               select(-NCHROBS, -CHR), 
                                             by = c("snps" = "SNP"))
# write out updated results file and put path in flanders input file
head(flanders_input)
dim(flanders_input)
write.table(flanders_input, input.info[1,"input"], sep="\t", quote=F, row.names=F)
read.table(input.info[1,"input"], nrows = 3, header = TRUE)

##### sdY ##### 
# get genes where eQTL were mapped because count matrix is subsetted by samples not genes
eQTL_genes_POPS2 <- unique(main_effects_pops$gene)
# calculate sd for each gene 
gene_sd_pops <- apply(gene_exp_pops, 1, sd)
# Make a data frame with gene IDs and SDs
gene_sd_df_pops <- data.frame(
  phenotype_id = names(gene_sd_pops),
  sdY = gene_sd_pops
) %>%
  filter(phenotype_id %in% eQTL_genes_POPS2)
# write out all sds in the correct format: TSV table with columns phenotype_id and sdY
gene_sd_df_pops %>% write.table(input.info[1,"sdY"], sep = "\t", quote = FALSE, row.names = FALSE)
read.table(input.info[1,"sdY"], nrows = 3, header = TRUE)

##### per-gene pval thresholds ##### 

# p-threhs1 come directly from eigenMT
pval_thresholds_pops %>% 
  select(-n.tests) %>%
  rename("gene" = "phenotype_id",
         "threshold" = "pval_thresh") %>%
  write.table(input.info[1,"per_gene_p1"], sep = "\t", quote = FALSE, row.names = FALSE)
read.table(input.info[1,"per_gene_p1"], nrows = 3, header = TRUE)
# p-thresh2 will come from p-thresh1 with caps
# if p1 < 1e-4:
#     p2 = min(10 * p1, 1e-4)
# elif p1 < 1e-3:
#     p2 = min(10 * p1, 1e-3)
# else:
#     p2 = min(5 * p1, 1e-2)
# For each gene, the variant inclusion threshold (p-thresh2) for Flanders fine-mapping was scaled from the gene-specific significance threshold (p-thresh1) estimated by eigenMT, reflecting the effective number of independent tests. We used a piecewise scaling with an upper cap: p-thresh2 = min(10 × p1, 1×10⁻⁴) for p1 < 1×10⁻⁴, p-thresh2 = min(10 × p1, 1×10⁻³) for 1×10⁻⁴ ≤ p1 < 1×10⁻³, and p-thresh2 = min(5 × p1, 1×10⁻²) for p1 ≥ 1×10⁻³. This approach allows inclusion of variants in LD with lead eQTLs while controlling region size, preserving stable fine-mapping performance.
pval_thresholds_pops %>% 
  mutate(p_thresh2 = ifelse(threshold < 1e-4, pmin(10*threshold, 1e-4), 
                            ifelse(threshold < 1e-3, pmin(10*threshold, 1e-3),
                                   pmin(5*threshold, 1e-2)))) %>%
    select(-n.tests, -threshold) %>%
  rename("gene" = "phenotype_id",
         "p_thresh2" = "pval_thresh") %>%
  write.table(input.info[1,"per_gene_p2"], sep = "\t", quote = FALSE, row.names = FALSE)
read.table(input.info[1,"per_gene_p2"], nrows = 3, header = TRUE)
```

# 2. Interval main effect eQTL from full cohort 
Include INTERVAL full cohort main effects in colocalization in order to determine whether the colocs with pregnancy traits could have been identified in any healthy cohort (with sufficient power)
Original publication: https://www.nature.com/articles/s41588-025-02096-3

Get list of genotyping sample IDs that had matched RNA-seq samples and were included in eQTL mapping 
```{r}
interval_metadata <- read.csv(paste0(interval_path, "recovery_intermediate_files/INTERVAL_RNAseq/expression_data/processed/INTERVAL_RNA_batch1-15_MasterCovariates_FilteredSamplesGenes_swapsSwappedMismatchRemoved.csv"))
dim(interval_metadata)  # 4732    1 # This matches the n samples in the publication 
# write out list of all genotyping IDs for this group of samples 
interval_metadata %>% mutate(fid = 0) %>%
  select(fid, affymetrix_ID) %>%
  write.table(paste0(interval_path, "POPS2_comparison/flanders_input/eQTL_genotyping_IDs.txt"), quote = FALSE, col.names = FALSE, row.names = FALSE, sep = "\t")
```

First, calculate frq on the genotyping data filtered to samples used for eQTL mapping  
```{bash}
cd interval_rna/POPS2_comparison/flanders_input
fash 20 
# load module with plink
module load HGI/softpack/groups/trynka/popgen/1.0
# calculate frq on the all autosomes file in only samples with RNA
plink \
  --bfile interval_rna/recovery_intermediate_files/INTERVAL_RNAseq/genotype/INTERVAL_RNAseq_Phase1-3_imputed_b38_biallelic_MAF0.005_AllAutosomes \
  --keep eQTL_genotyping_IDs.txt \
  --freq \
  --out INTERVAL_RNAseq_Phase1-3_imputed_b38_biallelic_MAF0.005_AllAutosomes_maf
  
```

Make input row
```{r}
input.info[2,"input"] <- paste0(interval_path, "POPS2_comparison/flanders_input/Interval_sumstats_forFlanders_sigEGenes.txt")
input.info[2,"per_gene_p1"] <- paste0(interval_path, "POPS2_comparison/flanders_input/interval_eqtl_pval_thresh1_per_gene_sigOnly.tsv")
input.info[2,"per_gene_p2"] <- paste0(interval_path, "POPS2_comparison/flanders_input/interval_eqtl_pval_thresh2_per_gene_sigOnly.tsv")
input.info[2, "rsid_lab"] <- "variant_id" # rsID or SNP identifier column name | String | Mandatory
input.info[2, "freq_lab"] <- "af_update" # Effect allele frequency column name | String or NA | Optional
input.info[2, "effect_lab"] <- "slope" # Effect size column (e.g., beta or OR) | String | Mandatory
input.info[2, "se_lab"] <- "slope_se" # Standard error column | String | Mandatory
input.info[2, "pvalue_lab"] <- "pval_nominal" # P-value column (nominal or log10) | String | Mandatory
input.info[2, "key"] <- 	"phenotype_id" # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[2, "type"] <- "quant" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[2, "is_molQTL"] <- 	TRUE
input.info[2, "grch"] <- 38 # Genome build (37 or 38) | Integer | Mandatory # From Interval eQTL paper: Reads were aligned to the GRCh38 human reference genome
input.info[2, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[2, "maf"] <- 	0.005 # Minor allele frequency threshold | Numeric (0–1) | Mandatory # From Interval methods: Before analysis, the dataset was restricted to individuals with RNA-seq and filtered to remove genetic variants with HWE exact test P<1e−6, genotype missingness >0.05 or minor allele frequency<0.5%.
input.info[2, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[2, "grch_bfile"] <- 38 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[2, "bfile"] <- "/path/to/dir/interval_rna/recovery_intermediate_files/INTERVAL_RNAseq/genotype/INTERVAL_RNAseq_Phase1-3_imputed_b38_biallelic_MAF0.005_AllAutosomes" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory # Since Interval main effects are only on autosomes, will use only autosome ref from interval 
input.info[2, "chr_lab"] <- "chr" # Chromosome column name in GWAS | String | Mandatory
input.info[2, "pos_lab"] <- "pos_b38" # Genomic position column name in GWAS | String | Mandatory
input.info[2, "a1_lab"] <- "effect_allele" # Effect allele column name | String | Mandatory 
input.info[2, "a0_lab"] <- "other_allele" # Non-effect allele column name | String | Mandatory 
input.info[2, "n_lab"] <- "nsamples" # Sample size column name | String or NA | Optional
input.info[2, "study_id"] <- "Interval_full_eQTL" # Identifier for the GWAS study | String | Mandatory
input.info[2,"p_thresh1"] <- 1.00E-05 # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory 
input.info[2,"p_thresh2"] <- 1.00E-04 # nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory 
input.info[2,"sdY"] <- 1 # For quantitative traits: population standard deviation of the trait (1 if standardised). If NA, will be approximated. NA for binary traits. | Numeric or NA | Mandatory # from Interval paper methods: Next, for each gene, the normalized log2-FPKM values across samples were transformed via the ranked-based inverse normal transformation function ‘rntransform’ implemented in the R package GenABEL (v1.8-0)59. Inverse normal transformation was applied to ensure the expression values followed a normal distribution.
input.info[2, "process_bfile"] <- TRUE
```


Then, make input files 
```{r}
# interval main effects are divided by chr
# first, list all of the results files 
interval_eQTL_res_files <- list.files(paste0(interval_path, "recovery_intermediate_files/INTERVAL_RNAseq/results_summary_stats/INTERVAL_eQTL_summary_statistics/"), 
                                      pattern = "INTERVAL_eQTL_nominal_chr*", 
                                      full.names = TRUE)
# read in the frq file
interval_frq <- fread(paste0(interval_path, "POPS2_comparison/flanders_input/INTERVAL_RNAseq_Phase1-3_imputed_b38_biallelic_MAF0.005_AllAutosomes_maf.frq"), header = TRUE)
# read in interval pval thresholds 
Interval_pval_thresholds <- read.delim(paste0(interval_path, "recovery_intermediate_files/INTERVAL_RNAseq/results_summary_stats/INTERVAL_eQTL_summary_statistics/INTERVAL_eQTL_phenotype_summary.tsv"), header = TRUE)

# make input files for flanders 
##### Results ##### 
# next read in all per-chr summary stats files 
interval_eQTL_list <- lapply(interval_eQTL_res_files, function(f) {
  fread(f, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
})
# combine them into one df
Interval_full_eQTL <- do.call(rbind, interval_eQTL_list)
# remove empty columns
Interval_full_eQTL <- Interval_full_eQTL %>% select(-V14, -V16)
# remove one row that looks like it must've been an error... slope_se == NA
Interval_full_eQTL <- Interval_full_eQTL %>% filter(!is.na(slope_se))
# join to maf calculated only in the data subset with RNA-seq samples 
# Add nsamples = 4732 to all rows as tensorQTL imputes genotypes 
Interval_full_eQTL <- Interval_full_eQTL %>% 
  # join to freq info 
  left_join(interval_frq, by = c("chr" = "CHR", "variant_id" = "SNP")) %>%
  # update AF based on the subsetted genotyping dataset
  mutate(af_update = ifelse(`A1` == `effect_allele`, MAF, 1-MAF),
         nsamples = 4732)
# remove eQTL results for genes that were determined not to be significant by tensorQTL
sig_eGenes <- Interval_pval_thresholds %>% filter(qval_sig == TRUE) %>% pull(phenotype_id) %>% unique()
Interval_sig_eQTL <- Interval_full_eQTL %>%
  filter(phenotype_id %in% sig_eGenes)

# write out updated results file and put path in flanders input file
write.table(Interval_sig_eQTL, input.info[2,"input"], sep="\t", quote=F, row.names=F)
read.table(input.info[2,"input"], nrows = 5, header = TRUE)

##### per-gene pval thresholds ##### 

# p-threhs1 come directly from tensorQTL output where pval_nominal_threshold is the threshold only for eGenes (qval_sig)
# otherwise, I will substitute a blanket pval threshold of 1e-5 (where the other option is to subset the input to only sig eQTL)
Interval_pval_thresholds %>% 
  filter(qval_sig == TRUE) %>% 
  rename("pval_nominal_threshold" = "pval_thresh") %>% 
  select(phenotype_id, pval_thresh) %>%
  write.table(input.info[2,"per_gene_p1"], sep = "\t", quote = FALSE, row.names = FALSE)
read.table(input.info[2,"per_gene_p1"], nrows = 3, header = TRUE)
# p-thresh2 will come from p-thresh1 with caps
# if p1 < 1e-4:
#     p2 = min(10 * p1, 1e-4)
# elif p1 < 1e-3:
#     p2 = min(10 * p1, 1e-3)
# else:
#     p2 = min(5 * p1, 1e-2)
# For each gene, the variant inclusion threshold (p-thresh2) for Flanders fine-mapping was scaled from the gene-specific significance threshold (p-thresh1) estimated by eigenMT, reflecting the effective number of independent tests. We used a piecewise scaling with an upper cap: p-thresh2 = min(10 × p1, 1×10⁻⁴) for p1 < 1×10⁻⁴, p-thresh2 = min(10 × p1, 1×10⁻³) for 1×10⁻⁴ ≤ p1 < 1×10⁻³, and p-thresh2 = min(5 × p1, 1×10⁻²) for p1 ≥ 1×10⁻³. This approach allows inclusion of variants in LD with lead eQTLs while controlling region size, preserving stable fine-mapping performance.
Interval_pval_thresholds %>% 
  filter(qval_sig == TRUE) %>% 
  rename("pval_nominal_threshold" = "pval_thresh") %>% 
  select(phenotype_id, pval_thresh) %>%
  mutate(p_thresh2 = ifelse(pval_thresh < 1e-4, pmin(10*pval_thresh, 1e-4), 
                            ifelse(pval_thresh < 1e-3, pmin(10*pval_thresh, 1e-3),
                                   pmin(5*pval_thresh, 1e-2)))) %>%
     select(phenotype_id, p_thresh2) %>%
  rename("p_thresh2" = "pval_thresh") %>%
  write.table(input.info[2,"per_gene_p2"], sep = "\t", quote = FALSE, row.names = FALSE)
read.table(input.info[2,"per_gene_p2"], nrows = 3, header = TRUE)
```

# 3. GDM GWAS
GDM GWAS EUR meta analysis from this preprint https://www.medrxiv.org/content/10.1101/2025.08.19.25333735v1.full (as recommended by Carolina)

Make input row
```{r}
# set parameters
input.info[3,"input"] <- paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/European_GDM_maternal.txt")
input.info[3, "rsid_lab"] <- "SNP" # rsID or SNP identifier column name | String | Mandatory
input.info[3, "freq_lab"] <- "FREQ" # Effect allele frequency column name | String or NA | Optional
input.info[3, "effect_lab"] <- "BETA" # Effect size column (e.g., beta or OR) | String | Mandatory
input.info[3, "se_lab"] <- "SE" # Standard error column | String | Mandatory
input.info[3, "pvalue_lab"] <- "PVAL" # P-value column (nominal or log10) | String | Mandatory
input.info[3, "key"] <- 	NA # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[3, "type"] <- "cc" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[3, "is_molQTL"] <- 	FALSE # | TRUE if molecular QTL (eQTL, pQTL); else FALSE | TRUE/FALSE | Mandatory
input.info[3, "grch"] <- 38 # Genome build (37 or 38) | Integer | Mandatory (from paper: If required, genetic data were transformed to align with the NCBI's Genome Reference Consortium human build 38 (GRCh38))
input.info[3, "p_thresh1"] <- 1e-05 #5e-08	#1.00E-05 # # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory # genome-wide significance threhsold 
input.info[3, "p_thresh2"] <- 1e-04 # 1e-05 # 1.00E-04 # Nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory # genome-wide suggestive significance threshold
input.info[3, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[3, "maf"] <- 	0.01 # Minor allele frequency threshold | Numeric (0–1) | Mandatory (from paper: After performing the multi-ancestry and ancestry-specific meta-analyses, we removed variants that were not present in at least half of the total meta-analysis sample size and variants with a MAF of less than 1%.)
input.info[3, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[3, "grch_bfile"] <- 38 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[3, "bfile"] <- "/path/to/dir/genotyping/data/imputed_genotyping_dataset_for_eQTL/POPS2_genotyping_imputed_eQTL_rsid" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory # Here I use the POPs2 genotyping dataset as an LD ref because it is a large, all female (X chr), majorly European cohort. X chr is coded as 23
input.info[3, "chr_lab"] <- "CHR" # Chromosome column name in GWAS | String | Mandatory
input.info[3, "pos_lab"] <- "BP" # Genomic position column name in GWAS | String | Mandatory
input.info[3, "a1_lab"] <- "A1" # Effect allele column name | String | Mandatory
input.info[3, "a0_lab"] <- "A2" # Non-effect allele column name | String | Mandatory
input.info[3, "n_lab"] <- "N" # Sample size column name | String or NA | Optional
input.info[3, "study_id"] <- "GDM_GWAS_EUR_meta" # Identifier for the GWAS study | String | Mandatory
input.info[3, "s"] <- 29640/(766250) # s | Proportion of cases (cc only) | Numeric (0–1) or NA | Mandatory # summary statistics are provided for Europeans (EUR: FG n = 17,810; 1hG n = 7,736; 2hG n = 10,784; HbA1c n = 5,377; GDM n = 766,250 [29,640 cases])
input.info[3, "process_bfile"] <- TRUE
input.info[3,"per_gene_p1"] <- ""#blank_pthresh1_filepath
input.info[3,"per_gene_p2"] <- ""#blank_pthresh2_filepath
```

Prepare input data 
```{r}
# read in gdm gwas
# downloaded from here https://rdm.uq.edu.au/files/88162382-7abb-4c07-a5bf-b3514b976602
gdm_gwas <- read.table(paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/776e0e38d862c1ce_GENDIP2-Q6147_Q6147/data/European_GDM_maternal"), header = TRUE)
# Note: No data from the X chr in these summary stats - I will follow up with Lizzy to see if this is expected. 

# write out the GWAS file as a tsv in the correct format 
write.table(gdm_gwas %>%
              mutate(A1 = toupper(A1),
                     A2 = toupper(A2)), 
            input.info[3,"input"], 
            sep="\t", quote=F, row.names=F) # make sure CHR is coded as 23
read.table(input.info[3,"input"], nrows = 3, header = TRUE)

```

# 4. Pre-eclampsia GWAS (FinnGen)
Website: https://r6.risteys.finngen.fi/phenocode/O15_PRE_OR_ECLAMPSIA and https://r12.finngen.fi/pheno/O15_PRE_OR_ECLAMPSIA (there are discrepancies, but the finngen website matches the downloaded metadata finngen_R12_pheno_n.tsv, so I will go with that n. Omar suggested it could have to do with some filtering.)
FinnGen summary stats already in humgen resources 

This dataset does appear to have X chr in the sumstats

```{bash}
# cd to correct dir
cd genotyping/analysis/eQTL/input_data/GWAS_summary_stats_for_coloc
# unzip sumstats from humgen resources to pops dir 
gunzip -c humgen/resources_v2/FinnGen/summary_stats/release/finngen_R12_O15_PREECLAMPS.gz > finngen_R12_O15_PREECLAMPS.txt
# remove the # at the beginning of the header row 
sed -i '1s/^#//' finngen_R12_O15_PREECLAMPS.txt
# get finnish reference from 1kgp
mkdir fin_ref
cd fin_ref
# get list of finnish samples (n = 105)
grep FIN humgen/resources_v2/1000g/release/20190312/20190312_biallelic_SNV_and_INDEL/IGSR_1000_Genomes_GRCh38_Samples.tsv | awk '{ print $1, $1 }' > 1000KG_FIN.txt
# load module
module load HGI/softpack/groups/trynka/popgen/1.0
# subset 1kgp
plink --bfile humgen/resources_v2/1000g/release/20190312/20190312_biallelic_SNV_and_INDEL/ALL.snp.GRCh38.phased.plink \
--set-missing-var-ids @_#_\$1_\$2 \
--keep 1000KG_FIN.txt \
--make-bed \
--out 1000G_FIN 
# get allele frequencies in this population 
plink --bfile 1000G_FIN --freq --out 1000G_FIN
```

Make Flanders input row
Info about summary stat columns here: https://finngen.gitbook.io/documentation/data-description
```{r}
# read in the head of the summary stats to understand columns
read.delim(
  paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/finngen_R12_O15_PREECLAMPS.txt"),
  nrows = 5
)

# set parameters
input.info[4,"input"] <- paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/finngen_R12_O15_PREECLAMPS.txt")
input.info[4, "rsid_lab"] <- "rsids" # rsID or SNP identifier column name | String | Mandatory # rsids = variant identifier
input.info[4, "freq_lab"] <- "af_alt" # Effect allele frequency column name | String or NA | Optional # af_alt = alternative (effect) allele frequency
input.info[4, "effect_lab"] <- "beta" # Effect size column (e.g., beta or OR) | String | Mandatory # beta = effect size (log(OR) scale) estimated with regenie for the alternative allele
input.info[4, "se_lab"] <- "sebeta" # Standard error column | String | Mandatory # sebeta = standard error of effect size estimated with regenie
input.info[4, "pvalue_lab"] <- "pval" # P-value column (nominal or log10) | String | Mandatory # pval = p-value from regenie
input.info[4, "key"] <- 	NA # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[4, "type"] <- "cc" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[4, "is_molQTL"] <- 	FALSE # | TRUE if molecular QTL (eQTL, pQTL); else FALSE | TRUE/FALSE | Mandatory
input.info[4, "grch"] <- 38 # Genome build (37 or 38) | Integer | Mandatory (from paper: If required, genetic data were transformed to align with the NCBI's Genome Reference Consortium human build 38 (GRCh38)) # https://docs.finngen.fi/finngen-data-specifics/red-library-data-individual-level-data/genotype-data/human-genome-build-we-use-and-how-do-we-report-genotypes Human reference genome build GRCh38/hg38 is used for FinnGen genotype data
input.info[4, "p_thresh1"] <- 1e-05 #5e-08	#1.00E-05 # # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory # genome-wide significance threhsold 
input.info[4, "p_thresh2"] <- 	1e-04 #1e-05 # 1.00E-04 # Nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory # genome-wide suggestive significance threshold
input.info[4, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[4, "maf"] <- 	3/(2*(9717+259181)) # Minor allele frequency threshold | Numeric (0–1) | Mandatory (from website: In sample-wise quality control steps, individuals with ambiguous gender, high genotype missingness (>5%), excess heterozygosity (+-4SD) and non-Finnish ancestry were excluded. In variant-wise quality control steps, variants with high missingness (>2%), low HWE P-value (<1e-6) and low minor allele count (MAC<3) were excluded.). I will therefore use 3/(2*n samples) or 3/(2*(n cases + n controls)) # 9717 cases, 259181 controls
input.info[4, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[4, "grch_bfile"] <- 38 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[4, "bfile"] <- "/path/to/dir/genotyping/analysis/eQTL/input_data/GWAS_summary_stats_for_coloc/fin_ref/1000G_FIN" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory 
input.info[4, "chr_lab"] <- "chrom" # Chromosome column name in GWAS | String | Mandatory # #chrom = chromosome on build GRCh38 (1-23)
input.info[4, "pos_lab"] <- "pos" # Genomic position column name in GWAS | String | Mandatory # pos = position in base pairs on build GRCh38
input.info[4, "a1_lab"] <- "alt" # Effect allele column name | String | Mandatory # alt = alternative allele (effect allele)
input.info[4, "a0_lab"] <- "ref" # Non-effect allele column name | String | Mandatory # ref = reference allele
input.info[4, "n_lab"] <- NA # Sample size column name | String or NA | Optional
input.info[4, "study_id"] <- "PE_FINNGEN_GWAS" # Identifier for the GWAS study | String | Mandatory
input.info[4, "s"] <- 9717/(9717+259181) # s | Proportion of cases (cc only) | Numeric (0–1) or NA | Mandatory # 9717 cases 259181 controls
input.info[4, "process_bfile"] <- TRUE
input.info[4,"per_gene_p1"] <- ""
input.info[4,"per_gene_p2"] <- ""
```

# 5. Pre-eclampsia GWAS (Meta analysis)

From Lizzy: pre-eclampsia Honigberg et al. 2023 (doi: 10.1038/s41591-023-02374-9) would actually be better-powered than the FinnGen one I shared before (we hadn't included it since it includes ALSPAC and MoBa).
Paper: https://www.nature.com/articles/s41591-023-02374-9#data-availability
Data download: https://figshare.com/articles/dataset/Polygenic_prediction_of_preeclampsia_and_gestational_hypertension/22680904/1

Make the input table
```{r}
read.table(paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/metal_preeclampsia_discovery_multiancestry_metaanalysis_formatted.txt"), nrows = 5, header = TRUE)

# set parameters
input.info[5,"input"] <- paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/metal_preeclampsia_discovery_multiancestry_metaanalysis_formatted.txt")
input.info[5, "rsid_lab"] <- "MarkerName" # rsID or SNP identifier column name | String | Mandatory # - this is the marker name
input.info[5, "freq_lab"] <- "Freq1" # Effect allele frequency column name | String or NA | Optional # - weighted average of frequency for allele 1 across all studies
input.info[5, "effect_lab"] <- "Effect" # Effect size column (e.g., beta or OR) | String | Mandatory # - overall estimated effect size for allele1
input.info[5, "se_lab"] <- "StdErr" # Standard error column | String | Mandatory # - overall standard error for effect size estimate
input.info[5, "pvalue_lab"] <- "P.value" # P-value column (nominal or log10) | String | Mandatory # - meta-analysis p-value
input.info[5, "key"] <- 	NA # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[5, "type"] <- "cc" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[5, "is_molQTL"] <- 	FALSE # | TRUE if molecular QTL (eQTL, pQTL); else FALSE | TRUE/FALSE | Mandatory
input.info[5, "grch"] <- 38 # Genome build (37 or 38) | Integer | Mandatory # Summary files provide information on chromosome, genomic position (NCBI build 37) # from paper: Variants from GWAS summary statistics were matched by genome build 38 position and alleles. GWAS summary statistics that were in genome build 37 were lifted to genome build 38 using UCSC liftOver
input.info[5, "p_thresh1"] <- 1e-05 #5e-08	#1.00E-05 # # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory # genome-wide significance threhsold 
input.info[5, "p_thresh2"] <- 	1e-04 #1e-05 # 1.00E-04 # Nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory # genome-wide suggestive significance threshold
input.info[5, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[5, "maf"] <- 	0.01 # Minor allele frequency threshold | Numeric (0–1) | Mandatory # from paper: We tested the association of common variants (minor allele frequency (MAF) ≥1%)
input.info[5, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[5, "grch_bfile"] <- 38 #37 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[5, "bfile"] <- "/path/to/dir/genotyping/data/imputed_genotyping_dataset_for_eQTL/POPS2_genotyping_imputed_eQTL_rsid" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory # Here I use the POPs2 genotyping dataset as an LD ref because it is a large, all female (X chr), majorly European cohort. X chr is coded as 23
input.info[5, "chr_lab"] <- "Chromosome" # Chromosome column name in GWAS | String | Mandatory
input.info[5, "pos_lab"] <- "Position" # Genomic position column name in GWAS | String | Mandatory
input.info[5, "a1_lab"] <- "Allele1" # Effect allele column name | String | Mandatory # - the first allele for this marker in the first file where it occurs
input.info[5, "a0_lab"] <- "Allele2" # Non-effect allele column name | String | Mandatory  # - the second allele for this marker in the first file where it occurs
input.info[5, "n_lab"] <- "NA" # Sample size column name | String or NA | Optional
input.info[5, "study_id"] <- "PE_GWAS_multiancestry_meta" # Identifier for the GWAS study | String | Mandatory
input.info[5, "s"] <- 20064/(20064+703117) # s | Proportion of cases (cc only) | Numeric (0–1) or NA | Mandatory # preeclampsia/eclampsia in 20,064 cases and 703,117 controls
input.info[5, "process_bfile"] <- TRUE
input.info[5,"per_gene_p1"] <- ""
input.info[5,"per_gene_p2"] <- ""
```

Format the sumstats (Alleles need to be uppercase)
```{r}
pe_multi_gwas <- read.table(paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/metal_preeclampsia_discovery_multiancestry_metaanalysis.txt"), 
                            header = TRUE, 
                            fill = TRUE)
pe_multi_gwas_formatted <- pe_multi_gwas %>%
              # remove NA values
              filter(if_all(everything(), ~ !is.na(.))) %>%
              # make the allele values uppercase 
              mutate(Allele1 = toupper(Allele1),
                     Allele2 = toupper(Allele2), 
                     # for the one row where chr == X, make chr = 23
                     Chromosome = ifelse(Chromosome == "X", 23, Chromosome))

# write out the GWAS file as a tsv in the correct format 
write.table(pe_multi_gwas_formatted, 
            input.info[5,"input"], 
            sep="\t", quote=F, row.names=F) # make sure CHR is coded as 23
read.table(input.info[5,"input"], nrows = 3, header = TRUE)
```



# 6. sPTB GWAS
Paper: https://pmc.ncbi.nlm.nih.gov/articles/PMC10101852/
Sumstat download: https://egg-consortium.org/Gestational-duration-2023.html (EGG_Maternal_GWAMA_allPTD.txt.gz. Maternal GWAS meta-analysis of preterm delivery in 233,290 women (n cases = 15,419). Data from 23andMe is not included in the meta-analysis.)
Downloaded to farm `genotyping/analysis/eQTL/input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_allPTD.tx`

This dataset does appear to have X chr in the sumstats

Make the input table
```{r}
# set parameters
input.info[6,"input"] <- paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_allPTD_formatted.txt")
input.info[6, "rsid_lab"] <- "RSID" # rsID or SNP identifier column name | String | Mandatory
input.info[6, "freq_lab"] <- "EAF" # Effect allele frequency column name | String or NA | Optional
input.info[6, "effect_lab"] <- "BETA" # Effect size column (e.g., beta or OR) | String | Mandatory
input.info[6, "se_lab"] <- "SE" # Standard error column | String | Mandatory
input.info[6, "pvalue_lab"] <- "PVALUE" # P-value column (nominal or log10) | String | Mandatory
input.info[6, "key"] <- 	NA # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[6, "type"] <- "cc" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[6, "is_molQTL"] <- 	FALSE # | TRUE if molecular QTL (eQTL, pQTL); else FALSE | TRUE/FALSE | Mandatory
input.info[6, "grch"] <- 37 # Genome build (37 or 38) | Integer | Mandatory # Summary files provide information on chromosome, genomic position (NCBI build 37)
input.info[6, "p_thresh1"] <- 1e-05 #5e-08	#1.00E-05 # # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory # genome-wide significance threhsold 
input.info[6, "p_thresh2"] <- 	1e-04 #1e-05 # 1.00E-04 # Nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory # genome-wide suggestive significance threshold
input.info[6, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[6, "maf"] <- 	0.01 # Minor allele frequency threshold | Numeric (0–1) | Mandatory # from paper: Genetic variants were excluded if genotype call rate <98%, Hardy-Weinberg equilibrium P value < 1 × 10−6 or minor allele frequency <1%.
input.info[6, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[6, "grch_bfile"] <- 38 #37 #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory
input.info[6, "bfile"] <- "/path/to/dir/genotyping/data/imputed_genotyping_dataset_for_eQTL/POPS2_genotyping_imputed_eQTL_rsid" #  Genome build of reference PLINK files (should match grch) | Integer (37 or 38) | Mandatory # Here I use the POPs2 genotyping dataset as an LD ref because it is a large, all female (X chr), majorly European cohort. X chr is coded as 23
input.info[6, "chr_lab"] <- "CHR_hg19" # Chromosome column name in GWAS | String | Mandatory
input.info[6, "pos_lab"] <- "POS_hg19" # Genomic position column name in GWAS | String | Mandatory
input.info[6, "a1_lab"] <- "EFF" # Effect allele column name | String | Mandatory
input.info[6, "a0_lab"] <- "REF" # Non-effect allele column name | String | Mandatory
input.info[6, "n_lab"] <- "TOTALSAMPLESIZE" # Sample size column name | String or NA | Optional
input.info[6, "study_id"] <- "sPTB_GWAS_EUR_META" # Identifier for the GWAS study | String | Mandatory
input.info[6, "s"] <- 15419/233290 # s | Proportion of cases (cc only) | Numeric (0–1) or NA | Mandatory # From download website (not paper, n samples differs because 23andme are excluded) Maternal GWAS meta-analysis of preterm delivery in 233,290 women (n cases = 15,419). Data from 23andMe is not included in the meta-analysis.
input.info[6, "process_bfile"] <- TRUE
input.info[6,"per_gene_p1"] <- ""
input.info[6,"per_gene_p2"] <- ""
```

Next, adjust the sumstats 
```{r}
# read them in... 
sPTB_GWAS_sumstats <- fread(
  paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_allPTD.txt"),
  header = TRUE,        # the first row is column names
  sep = "\t",           # TSV
  stringsAsFactors = FALSE,
  fill = TRUE           # fill missing columns with NA
)
# ensure your data.frame is a data.table
setDT(sPTB_GWAS_sumstats)
# fill empty rsids with something 
# fill missing RSID
sPTB_GWAS_sumstats[, RSID := fifelse(
  RSID == "" | is.na(RSID),
  paste0(CHR_hg19, "_", POS_hg19, "_", REF, "_", EFF),
  RSID
)]
# remove any row where REF and ALT aren't A, C, T or G (to get rid of the D/I deletion/insertion)
# define allowed alleles
valid_alleles <- c("A", "C", "T", "G")
# filter efficiently
sPTB_GWAS_sumstats <- sPTB_GWAS_sumstats[REF %in% valid_alleles & EFF %in% valid_alleles]
# this only filters out 6% of rows
# > 8863383/9348684
# [1] 0.9480888

# write this back out
write.table(sPTB_GWAS_sumstats, input.info[6,"input"], sep="\t", quote=F, row.names=F)
read.table(input.info[6,"input"], nrows = 3, header = TRUE)
```

# 7. Gestational duration GWAS
The sPTB GWAS paper states "We present a maternal genome-wide meta-analysis of gestational duration (n = 195,555), identifying 22 associated loci (24 independent variants) and an enrichment in genes differentially expressed during labor. A meta-analysis of preterm delivery (18,797 cases, 260,246 controls) revealed six associated loci and large genetic similarities with gestational duration.”
Because the sPTB sumstats have so few loci and this paper suggests that gestational duration is the "full" phenotype of preterm birth, I will also coloc with this set of sumstats.

Paper: https://pmc.ncbi.nlm.nih.gov/articles/PMC10101852/
Sumstat download: https://egg-consortium.org/Gestational-duration-2023.html (EGG_Maternal_GWAMA_GAraw.txt.gz. Maternal GWAS meta-analysis of gestational duration (days) in 151,987 women. Data from 23andMe is not included in the meta-analysis.)
Downloaded to farm `genotyping/analysis/eQTL/input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_allPTD.txt`

This dataset does appear to have X chr in the sumstats

Make the input table
```{r}
# set parameters
input.info[7,"input"] <- paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_GAraw_formatted.txt")
input.info[7, "rsid_lab"] <- "RSID" # rsID or SNP identifier column name | String | Mandatory
input.info[7, "freq_lab"] <- "EAF" # Effect allele frequency column name | String or NA | Optional
input.info[7, "effect_lab"] <- "BETA" # Effect size column (e.g., beta or OR) | String | Mandatory
input.info[7, "se_lab"] <- "SE" # Standard error column | String | Mandatory
input.info[7, "pvalue_lab"] <- "PVALUE" # P-value column (nominal or log10) | String | Mandatory
input.info[7, "key"] <- 	NA # Column identifying traits (e.g., gene ID) if is_molQTL=TRUE  | String or NA  | Conditional
input.info[7, "type"] <- "quant" # Trait type: 'quant' or 'cc' | quantitative trait (eQTL) or case control (GWAWS) | Mandatory 
input.info[7, "is_molQTL"] <- 	FALSE # | TRUE if molecular QTL (eQTL, pQTL); else FALSE | TRUE/FALSE | Mandatory
input.info[7, "grch"] <- 37 # Genome build (37 or 38) | Integer | Mandatory # Summary files provide information on chromosome, genomic position (NCBI build 37)
input.info[7, "p_thresh1"] <- 1e-05 #5e-08	#1.00E-05 # # nominal p-value threshold to retain significant association regions (suggested: 5e-08). | Numeric | Mandatory # genome-wide significance threhsold 
input.info[7, "p_thresh2"] <- 	1e-04 #1e-05 # 1.00E-04 # Nominal p-value threshold to define borders of association regions (suggested: 1e-05). | Numeric | Mandatory # genome-wide suggestive significance threshold
input.info[7, "hole"] <- 	200000 # Min distance to separate loci (e.g., 250000) | Integer | Mandatory # what is sensible
input.info[7, "maf"] <- 	0.01 # Minor allele frequency threshold | Numeric (0–1) | Mandatory # from paper: Genetic variants were excluded if genotype call rate <98%, Hardy-Weinberg equilibrium P value < 1 × 10−6 or minor allele frequency <1%.
input.info[7, "cs_thresh"] <- 	0.95 #  PIP threshold for credible sets (e.g., 0.99) | Numeric (0–1) | Mandatory # reduce to see if get fewer empty
input.info[7, "grch_bfile"] <- 38 #37 #   Genome build of LD reference PLINK files (should match grch if no liftover is used). | Integer (37 or 38) | Mandatory
input.info[7, "bfile"] <- "/path/to/dir/genotyping/data/imputed_genotyping_dataset_for_eQTL/POPS2_genotyping_imputed_eQTL_rsid" #  Path and file name prefix to PLINK binary files (.bed/.bim/.fam) used for LD computation | File path prefix (string) | Mandatory # Here I use the POPs2 genotyping dataset as an LD ref because it is a large, all female (X chr), majorly European cohort. X chr is coded as 23
input.info[7, "chr_lab"] <- "CHR_hg19" # Chromosome column name in GWAS | String | Mandatory
input.info[7, "pos_lab"] <- "POS_hg19" # Genomic position column name in GWAS | String | Mandatory
input.info[7, "a1_lab"] <- "EFF" # Effect allele column name | String | Mandatory
input.info[7, "a0_lab"] <- "REF" # Non-effect allele column name | String | Mandatory
input.info[7, "n_lab"] <- "TOTALSAMPLESIZE" # Sample size column name | String or NA | Optional
input.info[7, "study_id"] <- "GestDura_GWAS_EUR_META" # Identifier for the GWAS study | String | Mandatory
input.info[7, "s"] <- NA # s | Proportion of cases (cc only) | Numeric (0–1) or NA | Mandatory # This is a quantitative trait not a CC study 
input.info[7, "process_bfile"] <- TRUE
input.info[7,"per_gene_p1"] <- ""
input.info[7,"per_gene_p2"] <- ""
```

Adjust the sumstats 
```{r}
# read them in... 
GD_GWAS_sumstats <- fread(
  paste0(pops_lustre_path, "input_data/GWAS_summary_stats_for_coloc/EGG_Maternal_GWAMA_GAraw.txt"),
  header = TRUE,        # the first row is column names
  sep = "\t",           # TSV
  stringsAsFactors = FALSE,
  fill = TRUE           # fill missing columns with NA
)
# ensure your data.frame is a data.table
setDT(GD_GWAS_sumstats)
# remove any row where REF and ALT aren't A, C, T or G (to get rid of the D/I deletion/insertion)
# define allowed alleles
valid_alleles <- c("A", "C", "T", "G")
# filter efficiently
GD_GWAS_sumstats <- GD_GWAS_sumstats[REF %in% valid_alleles & EFF %in% valid_alleles]
# this only filters out 4% of rows
# > 9165126/9471983
# [1] 0.9676037

# write this back out
write.table(GD_GWAS_sumstats, input.info[7,"input"], sep="\t", quote=F, row.names=F)
read.table(input.info[7,"input"], nrows = 3, header = TRUE)
```

# Finally, write out flanders input table
```{r}
write.table(input.info, 
            paste0(outpath, "POPS2_flanders_input.txt"), 
            sep="\t", row.names=F, quote=F)
```

# Run flanders

Run the Flanders pipeline for finemapping and colocalization developed by HT (https://github.com/Biostatistics-Unit-HT/Flanders)
However, we run this version updated to allow per-gene p-value thresholds by HGI:  
github.com/wtsi-hgi/Flanders/tree/per-gene-p-value 

```{bash}
module load HGI/pipelines/flanders/v2026_01_15
# cd to the correct directory 
genotyping/analysis/eQTL/output_data/flanders_5
# run flanders
flanders --summarystats_input ../../input_data/flanders/POPS2_flanders_input.txt --run_liftover TRUE --susie_max_iter 1000 --pph4_threshold 0.8 --pph3_threshold 0.8 --chromosomes 1-23 -config ../../input_data/flanders/flanders_eQTL_nextflow4.config --publish_susie TRUE
```